---
title: 'tidyTuesday: coffee chains'
author: 'Robert W. Walker'
date: '2018-05-09'
categories:
  - R
  - tidyverse
  - Maps
tags:
  - R
  - tidyTuesday
  - tidyverse
bibliography: "bibliography.bib"
nocite: |
     @*
format: 
   html:
     code-fold: true
     code-copy: hover
     code-block-border-left: true
     df-print: paged
     fig-format: retina
     footnotes-hover: true
     html-math-method: katex
     toc: true
     toc-title: Navigation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, message=FALSE, warning=FALSE)
```


# The tidyTuesday for this week is coffee chain locations

For this week:
1. The basic link to the [`#tidyTuesday`](https://github.com/rfordatascience/tidytuesday) shows [an original article](http://flowingdata.com/2014/03/18/coffee-place-geography/) for Week 6.

The page notes that Starbucks, Tim Horton, and Dunkin Donuts have raw data available.

```{r LoadData, message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(janitor)
library(geofacet)
library(ggbeeswarm)
library(ggrepel)
Starbucks <- readRDS(url("https://github.com/robertwwalker/academic-mymod/raw/master/data/week6SB.rds"))
Dunkin.Donuts <- readRDS(url("https://github.com/robertwwalker/academic-mymod/raw/master/data/week6DD.rds"))
Tim.Hortons <- readRDS(url("https://github.com/robertwwalker/academic-mymod/raw/master/data/week6TH.rds"))
```

What do the data look like?

### Starbucks Data

```{r}
library(skimr)
skim(Starbucks)
```

### Dunkin Donuts Data


```{r}
skim(Dunkin.Donuts)
```

### Tim Horton's Data

```{r}
skim(Tim.Hortons)
```

## Plot Starbucks

A basic plot of the global Starbucks data.

```{r}
library(ggmap)
mapWorld <- borders("world", colour="gray50", fill="gray50") # create a layer of borders
mp <- ggplot() +   mapWorld
mp <- mp + geom_point(aes(x=Starbucks$Longitude, y=Starbucks$Latitude) ,color="dark green", size=0.5) + xlab("") + ylab("")
mp <- mp + geom_point(aes(x=Dunkin.Donuts$loc_LONG_centroid, y=Dunkin.Donuts$loc_LAT_centroid) ,color="orange", size=0.5) + xlab("") + ylab("") + theme_void()
mp
```

## Starbucks and Dunkin

Google Maps interface changed and I have not updated this part.  Shame.

```{r, eval=FALSE}
coffee.xl$timhorton <- coffee.xl$timhorton %>% mutate(full.address=paste(address,city,state,toupper(country),sep=", "))
# Tim.H <- coffee.xl$timhorton %>% filter(country=="us") %>% mutate_geocode(full.address, source="dsk")
load(paste0(here(),"/data/geocoded.RData"))
THUS <- Tim.H %>% select(country,full.address,lon,lat)
names(THUS) <- c("Country","Address","Longitude","Latitude")
THUS$chain <- "Tim Horton's"
DDUS <- coffee.xl$dunkin %>% select(e_country, e_address, loc_LONG_centroid,loc_LAT_centroid)
names(DDUS) <- c("Country","Address","Longitude","Latitude")
DDUS$chain <- "Dunkin"
SBUX <- coffee.xl$starbucks %>% filter(Country=="us") %>% select(Country,`Street Address`,Longitude,Latitude)
names(SBUX) <- c("Country","Address","Longitude","Latitude")
US.Map.Coffee <- bind_rows(DDUS,SBUX,THUS)
USmap <- get_stamenmap(us, zoom = 5)
mp <- ggmap(USmap)
mp <- mp + geom_point(aes(x=Starbucks$Longitude, y=Starbucks$Latitude) ,color="dark green", size=0.5) + xlab("") + ylab("")
mp <- mp + geom_point(aes(x=Dunkin.Donuts$loc_LONG_centroid, y=Dunkin.Donuts$loc_LAT_centroid) ,color="orange", size=0.5) + xlab("") + ylab("")
mp
```


# References

```{r}
knitr::write_bib(names(sessionInfo()$otherPkgs), file="bibliography.bib")
```

